### 2018-11-21 manuscript-final.tex
final ms

% Use the PLoS provided BiBTeX style
\bibliographystyle{plos2015}
%\bibliographystyle{apalike}

@@

\section*{Discussion}
The framework  we presented in this article describes a general computational set of tools to store, manipulate and analyse behavioural data.
We showed that \texttt{rethomics} can readily be used to analyse circadian data. As such, it already fills a clear gap in the sleep and circadian community by providing an open, extensible and programmatic toolbox as opposed to a rigid array of GUI tools which cannot interoperate.
In particular, within the \texttt{R} environment, \texttt{rethomic} features a streamlined workflow (Fig.~\ref{fig:fig-1}):
data loading; quality control (Fig.~\ref{fig:fig-3}); per individual visualisation (Fig.~\ref{fig:fig-4}A) and computation (Fig.~\ref{fig:fig-4}B); group averages and statistical analysis (Fig.~\ref{fig:fig-5}), ultimately resulting in publication-quality figures.
In addition to the canonical analysis we presented in this study, the richness of the \texttt{R} environment renders rethomics particularly interesting for exploratory work such as the use of continuous wavelet transform to study the modulation of walking pace(Fig.~\ref{fig:fig-6}).

Altogether, our tool provides a needed opportunity for the community to share analytic and visualisation solutions whilst embracing the diversity of recording equipment. By design, rethomics is very flexible and not limited to the study of circadian rhythms.
In fact, \texttt{behavr} tables (Fig.~\ref{fig:fig-2}) are a general structure to link data to metadata that come with no assumption regarding the type and number of variables stored.
Therefore, we anticipate it will apply to  a variety of contexts and scales.
For instance, in the last decade, multiple devices have been developed -- either commercially or academically -- to record longitudinal data on several animal models (\emph{i.e.} feeding, metabolic activity, heart rate, position and so on). In collaboration with the teams providing of these methods, we envisage the development of device-specific packages that will load data into \texttt{rethomics}, making visualisation and analysis immediately available. 
For instance, the developer any behaviour recording method can simply write a small package to read their data in a \texttt{behavr} table and benefit from the functionalities available.
Furthermore, such an approach becomes particularily useful in order to compare methods or study behaviours that are recorded by different devices, on the same individuals. Indeed, in these contexts, it becomes apparent that analysis must be unified at some level.

In conclusion, \texttt{rethomics} already delivers state-of-the-art tools to study circadian rhythm using conventional tools, but was designed in an inclusive and extensible manner. As such, if fills a unique niche in a community that is becoming increasingly quantitative.

\section*{Availability}
All \texttt{rethomics} packages are listed at \href{https://rethomics.github.io/intro.html}{https://rethomics.github.io/intro.html}.
and are hosted on the official Comprehensive \texttt{R} Archive Network (CRAN) under the terms of the GPLv3 license. 
Extensive installation instructions, as well as reproducible demos and tutorials, are available at
\href{https://rethomics.github.io/}{https://rethomics.github.io/}.
All packages are continuously integrated and unit tested on several versions of \texttt{R} to minimise the risk of present and future issues.

@@

\bibliography{manuscript}{}

>>>>

% Use the PLoS provided BiBTeX style
%\bibliographystyle{plos2015}
%\bibliographystyle{apalike}

@@

\section*{Discussion}

The framework we presented in this article describes a general computational set of tools to store, manipulate and analyse behavioural data. 
As a practical case study, in this article, we showed how \texttt{rethomics} can readily be used to analyse circadian and sleep data.
As such, it already fills a clear gap in the sleep and circadian community by providing an open, extensible and programmatic toolbox as opposed to a rigid array of GUI tools which cannot interoperate. 
In particular, within the R environment, rethomic features a streamlined workflow (Fig.~\ref{fig:fig-1}): data loading; quality control (Fig.~\ref{fig:fig-3}); per-individual visualisation (Fig.~\ref{fig:fig-4}A) and computation (Fig.~\ref{fig:fig-4}B); group averages and statistical analysis (Fig.~\ref{fig:fig-5}), ultimately resulting in publication-quality figures. 
In addition to the canonical analysis we presented in this study, the richness of the \texttt{R} environment makes \texttt{rethomics} particularly interesting for exploratory work: as a case in point, we showed here the use of continuous wavelet transform to study the modulation of walking pace (Fig.~\ref{fig:fig-6}). Altogether, \texttt{rethomics} will allow the community to share solutions for the analysis and visualisation of behavioural data originated from a plethora of recording equipment. 
By design, \texttt{rethomics} is meant to be flexible and certainly not limited to the study of circadian rhythms.
In fact, \texttt{behavr} tables (Fig.~\ref{fig:fig-2}) are a general structure to link data to metadata that come with no assumption regarding the type and number of variables stored.
Therefore, we anticipate it will apply to a variety of contexts and scales.
In the last decade, multiple devices have been developed -- either commercially or academically -- to record longitudinal data on several animal models (i.e. feeding, metabolic activity, heart rate, position and so on).
All these devices and behavioural paradigm could take advantage of the skeletal structure in \texttt{rethomics} and lead to  the development of device-specific packages.
Such an approach becomes particularly useful in order to compare methods or study behaviours that are recorded by different devices on the same individual animals.

\section*{Availability}
All \texttt{rethomics} packages are listed at \href{https://rethomics.github.io/intro.html}{https://rethomics.github.io/intro.html}.
and are hosted on the official Comprehensive \texttt{R} Archive Network (CRAN) under the terms of the GPLv3 license. 
Installation instructions, as well as reproducible demos and tutorials, are available at
\href{https://rethomics.github.io/}{https://rethomics.github.io/}.
All packages are continuously integrated and unit tested on several versions of \texttt{R} to minimise the risk of present and future issues.

@@

\begin{thebibliography}{10}

	\bibitem{sokolowski_drosophila_2001}
	Sokolowski MB.
	\newblock {\emph{Drosophila}}: {{Genetics}} Meets Behaviour.
	\newblock Nature Reviews Genetics. 2001;2(11):879--890.
	\newblock doi:{10.1038/35098592}.

	\bibitem{stephens_big_2015}
	Stephens ZD, Lee SY, Faghri F, Campbell RH, Zhai C, Efron MJ, et~al.
	\newblock Big {{Data}}: {{Astronomical}} or {{Genomical}}?
	\newblock PLOS Biology. 2015;13(7):e1002195.
	\newblock doi:{10.1371/journal.pbio.1002195}.

	\bibitem{schatz_biological_2015}
	Schatz MC.
	\newblock Biological Data Sciences in Genome Research.
	\newblock Genome Research. 2015;25(10):1417--1422.
	\newblock doi:{10.1101/gr.191684.115}.

	\bibitem{sejnowski_putting_2014}
	Sejnowski TJ, Churchland PS, Movshon JA.
	\newblock {Putting big data to good use in neuroscience}.
	\newblock Nature Neuroscience. 2014;17(11):1440--1441.
	\newblock doi:{10.1038/nn.3839}.

	\bibitem{branson_high-throughput_2009}
	Branson K, Robie AA, Bender J, Perona P, Dickinson MH.
	\newblock High-Throughput Ethomics in Large Groups of {{{\emph{Drosophila}}}}.
	\newblock Nature Methods. 2009;6(6):451--457.
	\newblock doi:{10.1038/nmeth.1328}.

	\bibitem{reiser_ethomics_2009}
	Reiser M.
	\newblock The Ethomics Era?
	\newblock Nature Methods. 2009;6(6):413--414.
	\newblock doi:{10.1038/nmeth0609-413}.

	\bibitem{anderson_toward_2014}
	Anderson DJ, Perona P.
	\newblock Toward a {{Science}} of {{Computational Ethology}}.
	\newblock Neuron. 2014;84(1):18--31.
	\newblock doi:{10.1016/j.neuron.2014.09.005}.

	\bibitem{brown_ethology_2018}
	Brown AEX, de~Bivort B.
	\newblock Ethology as a Physical Science.
	\newblock Nature Physics. 2018; p.~1.
	\newblock doi:{10.1038/s41567-018-0093-0}.

	\bibitem{berman_measuring_2018}
	Berman GJ.
	\newblock Measuring Behavior across Scales.
	\newblock BMC Biology. 2018;16:23.
	\newblock doi:{10.1186/s12915-018-0494-7}.

	\bibitem{faville_how_2015}
	Faville R, Kottler B, Goodhill GJ, Shaw PJ, {van Swinderen} B.
	\newblock How Deeply Does Your Mutant Sleep? {{Probing}} Arousal to Better
	Understand Sleep Defects in {{{\emph{Drosophila}}}}.
	\newblock Scientific Reports. 2015;5:8454.
	\newblock doi:{10.1038/srep08454}.

	\bibitem{pelkowski_novel_2011}
	Pelkowski SD, Kapoor M, Richendrfer HA, Wang X, Colwill RM, Creton R.
	\newblock A Novel High-Throughput Imaging System for Automated Analyses of
	Avoidance Behavior in Zebrafish Larvae.
	\newblock Behavioural Brain Research. 2011;223(1):135--144.
	\newblock doi:{10.1016/j.bbr.2011.04.033}.

	\bibitem{itskov_automated_2014}
	Itskov PM, Moreira JM, Vinnik E, Lopes G, Safarik S, Dickinson MH, et~al.
	\newblock Automated Monitoring and Quantitative Analysis of Feeding Behaviour
	in {{{\emph{Drosophila}}}}.
	\newblock Nature Communications. 2014;5:4560.
	\newblock doi:{10.1038/ncomms5560}.

	\bibitem{ro_flic_2014}
	Ro J, Harvanek ZM, Pletcher SD.
	\newblock {{FLIC}}: {{High}}-{{Throughput}}, {{Continuous Analysis}} of
	{{Feeding Behaviors}} in {{Drosophila}}.
	\newblock PLOS ONE. 2014;9(6):e101107.
	\newblock doi:{10.1371/journal.pone.0101107}.

	\bibitem{dankert_automated_2009}
	Dankert H, Wang L, Hoopfer ED, Anderson DJ, Perona P.
	\newblock Automated {{Monitoring}} and {{Analysis}} of {{Social Behavior}} in
	{{Drosophila}}.
	\newblock Nature methods. 2009;6(4):297--303.
	\newblock doi:{10.1038/nmeth.1310}.

	\bibitem{tsai_image_2012}
	Tsai HY, Huang YW.
	\newblock Image {{Tracking Study}} on {{Courtship Behavior}} of {{Drosophila}}.
	\newblock PLOS ONE. 2012;7(4):e34784.
	\newblock doi:{10.1371/journal.pone.0034784}.

	\bibitem{swierczek_high-throughput_2011}
	Swierczek NA, Giles AC, Rankin CH, Kerr RA.
	\newblock High-Throughput Behavioral Analysis in {{{\emph{C}}}}{\emph{.
			Elegans}}.
	\newblock Nature Methods. 2011;8(7):592--598.
	\newblock doi:{10.1038/nmeth.1625}.

	\bibitem{perez-escudero_idtracker_2014}
	P{\'e}rez-Escudero A, Vicente-Page J, Hinz RC, Arganda S, {de Polavieja} GG.
	\newblock {{idTracker}}: Tracking Individuals in a Group by Automatic
	Identification of Unmarked Animals.
	\newblock Nature Methods. 2014;11(7):743--748.
	\newblock doi:{10.1038/nmeth.2994}.

	\bibitem{robie_mapping_2017}
	Robie AA, Hirokawa J, Edwards AW, Umayam LA, Lee A, Phillips ML, et~al.
	\newblock Mapping the {{Neural Substrates}} of {{Behavior}}.
	\newblock Cell. 2017;170(2):393--406.e28.
	\newblock doi:{10.1016/j.cell.2017.06.032}.

	\bibitem{kabra_jaaba_2013}
	Kabra M, Robie AA, Rivera-Alba M, Branson S, Branson K.
	\newblock {{JAABA}}: Interactive Machine Learning for Automatic Annotation of
	Animal Behavior.
	\newblock Nature Methods. 2013;10(1):64--67.
	\newblock doi:{10.1038/nmeth.2281}.

	\bibitem{lopes_bonsai_2015}
	Lopes G, Bonacchi N, Fraz{\~a}o J, Neto JP, Atallah BV, Soares S, et~al.
	\newblock Bonsai: An Event-Based Framework for Processing and Controlling Data
	Streams.
	\newblock Frontiers in Neuroinformatics. 2015;9.
	\newblock doi:{10.3389/fninf.2015.00007}.

	\bibitem{geissmann_ethoscopes_2017}
	Geissmann Q, Rodriguez LG, Beckwith EJ, French AS, Jamasb AR, Gilestro GF.
	\newblock Ethoscopes: {{An}} Open Platform for High-Throughput Ethomics.
	\newblock PLOS Biology. 2017;15(10):e2003026.
	\newblock doi:{10.1371/journal.pbio.2003026}.

	\bibitem{roumpeka_review_2017}
	Roumpeka DD, Wallace RJ, Escalettes F, Fotheringham I, Watson M.
	\newblock A {{Review}} of {{Bioinformatics Tools}} for {{Bio}}-{{Prospecting}}
	from {{Metagenomic Sequence Data}}.
	\newblock Frontiers in Genetics. 2017;8.
	\newblock doi:{10.3389/fgene.2017.00023}.

	\bibitem{huber_orchestrating_2015}
	Huber W, Carey VJ, Gentleman R, Anders S, Carlson M, Carvalho BS, et~al.
	\newblock Orchestrating High-Throughput Genomic Analysis with {{Bioconductor}}.
	\newblock Nature Methods. 2015;12(2):115--121.
	\newblock doi:{10.1038/nmeth.3252}.

	\bibitem{leipzig_review_2017}
	Leipzig J.
	\newblock A Review of Bioinformatic Pipeline Frameworks.
	\newblock Briefings in Bioinformatics. 2017;18(3):530--536.
	\newblock doi:{10.1093/bib/bbw020}.

	\bibitem{peng_reproducible_2011}
	Peng RD.
	\newblock Reproducible {{Research}} in {{Computational Science}}.
	\newblock Science. 2011;334(6060):1226--1227.
	\newblock doi:{10.1126/science.1213847}.

	\bibitem{stodden_toward_2013}
	Stodden V, Guo P, Ma Z.
	\newblock Toward {{Reproducible Computational Research}}: {{An Empirical
			Analysis}} of {{Data}} and {{Code Policy Adoption}} by {{Journals}}.
	\newblock PLOS ONE. 2013;8(6):e67111.
	\newblock doi:{10.1371/journal.pone.0067111}.

	\bibitem{hashem_rise_2015}
	Hashem IAT, Yaqoob I, Anuar NB, Mokhtar S, Gani A, Ullah~Khan S.
	\newblock The Rise of ``Big Data'' on Cloud Computing: {{Review}} and Open
	Research Issues.
	\newblock Information Systems. 2015;47:98--115.
	\newblock doi:{10.1016/j.is.2014.07.006}.

	\bibitem{r_core_team_r_2017}
	{R Core Team}.
	\newblock R: {{A Language}} and {{Environment}} for {{Statistical Computing}}.
	\newblock Vienna, Austria: {R Foundation for Statistical Computing}; 2017.

	\bibitem{wickham_tidyverse_2017}
	Wickham H.
	\newblock Tidyverse: {{Easily Install}} and {{Load}} the '{{Tidyverse}}'; 2017.

	\bibitem{dowle_data.table_2017}
	Dowle M, Srinivasan A.
	\newblock Data.Table: {{Extension}} of `data.Frame`; 2017.

	\bibitem{wickham_ggplot2_2016}
	Wickham H.
	\newblock Ggplot2: {{Elegant Graphics}} for {{Data Analysis}}.
	\newblock {Springer}; 2016.

	\bibitem{sokolove_chi_1978}
	Sokolove PG, Bushell WN.
	\newblock The Chi Square Periodogram: {{Its}} Utility for Analysis of Circadian
	Rhythms.
	\newblock Journal of Theoretical Biology. 1978;72(1):131--160.
	\newblock doi:{10.1016/0022-5193(78)90022-X}.

	\bibitem{ruf_lomb-scargle_1999}
	Ruf T.
	\newblock The {{Lomb}}-{{Scargle Periodogram}} in {{Biological Rhythm
			Research}}: {{Analysis}} of {{Incomplete}} and {{Unequally Spaced
			Time}}-{{Series}}.
	\newblock Biological Rhythm Research. 1999;30(2):178--201.
	\newblock doi:{10.1076/brhm.30.2.178.1422}.

	\bibitem{dubowy_circadian_2017}
	Dubowy C, Sehgal A.
	\newblock Circadian {{Rhythms}} and {{Sleep}} in {{Drosophila}} Melanogaster.
	\newblock Genetics. 2017;205(4):1373--1397.
	\newblock doi:{10.1534/genetics.115.185157}.

	\bibitem{buhl_quasimodo_2016}
	Buhl E, Bradlaugh A, Ogueta M, Chen KF, Stanewsky R, Hodge JJL.
	\newblock Quasimodo Mediates Daily and Acute Light Effects on {{Drosophila}}
	Clock Neuron Excitability.
	\newblock Proceedings of the National Academy of Sciences.
	2016;113(47):13486--13491.
	\newblock doi:{10.1073/pnas.1606547113}.

	\bibitem{ogueta_maite_2018_1172980}
	Ogueta M, Stanewsky R.
	\newblock {{LL Behaviour}} of {{TIM Gal4}} $>$ {{NKCC OX}} and {{NKCC OX}} / +
	Flies.
	\newblock Zenodo. 2018;doi:{10.5281/zenodo.1172980}.

	\bibitem{aguiar-conraria_business_2011}
	Aguiar-Conraria L, Joana~Soares M.
	\newblock Business Cycle Synchronization and the {{Euro}}: {{A}} Wavelet
	Analysis.
	\newblock Journal of Macroeconomics. 2011;33(3):477--489.
	\newblock doi:{10.1016/j.jmacro.2011.02.005}.

	\bibitem{lau_climate_1995}
	Lau KM, Weng H.
	\newblock Climate {{Signal Detection Using Wavelet Transform}}: {{How}} to
	{{Make}} a {{Time Series Sing}}.
	\newblock Bulletin of the American Meteorological Society.
	1995;76(12):2391--2402.
	\newblock doi:{10.1175/1520-0477(1995)076<2391:CSDUWT>2.0.CO;2}.

	\bibitem{cazelles_wavelet_2008}
	Cazelles B, Chavez M, Berteaux D, M{\'e}nard F, Vik JO, Jenouvrier S, et~al.
	\newblock Wavelet Analysis of Ecological Time Series.
	\newblock Oecologia. 2008;156(2):287--304.
	\newblock doi:{10.1007/s00442-008-0993-2}.

	\bibitem{grossmann_decomposition_1984}
	Grossmann A, Morlet J.
	\newblock Decomposition of {{Hardy Functions}} into {{Square Integrable
			Wavelets}} of {{Constant Shape}}.
	\newblock SIAM Journal on Mathematical Analysis. 1984;15(4):723--736.
	\newblock doi:{10.1137/0515056}.

	\bibitem{leise_wavelet_2013}
	Leise TL.
	\newblock Wavelet Analysis of Circadian and Ultradian Behavioral Rhythms.
	\newblock Journal of Circadian Rhythms. 2013;11:5.
	\newblock doi:{10.1186/1740-3391-11-5}.

	\bibitem{schmidbauer_waveletcomp_2018}
	Schmidbauer AR, Harald. {{WaveletComp}}: {{Computational Wavelet Analysis}};
	2018.

\end{thebibliography}
